\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2024-02-09 RefereeResponse/response/RefereeResponse.tex

\usepackage{babel}
\begin{document}

\title{Response to Editor: JHEP\_031P\_1223}

\date{\today}

\maketitle
Dear Editor,

We would like to begin by sincerely thanking the referee for their careful reading of the manuscript.
We believe their thoughtful comments and suggestions have produced a better paper.
The referee suggested more context for the two-loop results, and noted that a comparison with $\mathcal{N}=4$ super-Yang Mills (sYM) was absent. We agree that adding this context provides a necessary framing of the main technical results.
To this end, below is a summary of the modifications to the manuscript:
\begin{itemize}
\item A small comment has been added to the introduction at the top of page 2 reminding readers that color-dual representations of sYM have been found to high loop order.
\item At the bottom of page 2, when stating that there is no local 4pt 2-loop numerator for pure YM we remind readers that such a numerator exists for sYM.
\item At the end of the introduction at the top of page 3 we clarify that the bowtie cut failure is specific to pure YM.
We go on to speculate that this is because sYM and NLSM are simpler theories, at least as far as color-kinematics is concerned.
\item In addition to the comments in the introduction and section 4, we have included one additional modification.
Section 2.3 deals with ansatz locality so this seemed like a natural place to mention that the sYM ans\"atze incorporate locality as a key simplifying assumption.
At the bottom of page 7 and the top of page 8, we emphasize that the sYM numerators through four loops are local.
We comment in footnote 4 that the full set of local numerators has never been fully explored for sYM at 4pt 5-loop.
\item The first paragraph in section 4 closes by putting our results in the context of sYM, specifically by reminding readers of the state of the art in that theory. Another such reminder is included in footnote 10 at the beginning of section 4.3. 
\item After stating the bowtie failure condition for pure YM in (4.9) we give specific reasons for why NLSM and sYM are more privileged than pure YM consistent with the ``bow-tie'' cut. 
\end{itemize}

We would like to thank the referees again for their carefully considered feedback.
Now that the referees comments have been addressed, and we believe that the improved the readability of the manuscript makes it an excellent fit for JHEP.

### 2024-02-09 RefereeResponse/paper/2loopNLSM.tex
added commentary, edited response

\abstract{We report on progress in understanding how to construct color-dual multi-loop amplitudes.
First we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators through one-loop directly from Feynman rules. By a simple weight counting argument, this approach does not further generalize to two-loops.  As a first step in
  understanding the two-loop challenge, we use a $D$-dimensional color-dual bootstrap to 
  successfully construct globally color-dual
  \emph{local} two-loop four-point nonlinear sigma model (NLSM) numerators.  The double-copy of
these NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known unitarity constructed integrands 
  of special Galileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively. 
  Applying our bootstrap to two-loop four-point pure Yang-Mills, we exhaustively search the space of local numerators and find that it fails to satisfy global color-kinematics duality, completing a search previously initiated in the literature. 
  We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

@@

\section{Introduction}\label{intro}
Since the turn of the century, our understanding of the $S$-matrix and
its concealed structures has expanded dramatically. At the heart of this
progress is the mantra that physical observables are simpler when
studied on-shell
\cite{Parke:1986gb,TasiLance,Cheung:2017pzi}. However, while inserting on-shell states certainly offers a path towards taming the
factorial growth of Feynman diagrams, it also obscures the off-shell
simplicity at the heart of many quantum field theories. A shining
example of this hidden structure is the duality between color and
kinematics \cite{BCJ,Bern:2010ue,BCJreview}, which states that the
kinematic numerators of gauge theory amplitudes can be rearranged to
obey the same algebraic identities as the constituent color
factors. When this duality is realized globally off-shell, it
dramatically decreases the combinatorial complexity posed by integrand
construction, and offers a path towards efficient assembly of quantum
gravity integrands directly from simpler gauge theory building blocks
via the double copy \cite{BCJ,Bern:2010ue}.

@@

Many of the one-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed.  Since the sum only contains $(n-2)!$
terms, there must be additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental Bern-Carrasco-Johansson (BCJ) identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half-ladder because any
diagram can be reduced to a half-ladder through successive
applications of the Jacobi identity.  At one-loop, every diagram can
be related to one topology as well, in this case the $n$-gon, which
can be understood as a forward limit of the half-ladder.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory. We
refer the reader to Ref.~\cite{BCJreview} for more background on the
topic.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to trivialize
the duality by encoding it in an off-shell Lagrangian
of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}\label{cubicCKLag}
\mathcal{L} = P_V^{ab}P_W^{\mu\nu}\mathcal{O}^a_\mu \Box \mathcal{O}^b_\nu +  V^{abc} W^{\mu \nu \rho}\mathcal{O}^a_\mu \mathcal{O}^b_\nu \mathcal{O}^c_\rho
\end{equation}
where $V^{abc}$ and $W^{\mu \nu \rho}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_\mu$ indexed by quantum numbers $a$ and
$\mu$. The Feynman rules for this theory are simply:
\be\label{CKFrules}
\begin{split}
\propFrule{$a,\mu$}{$b,\nu$}{$k \rightarrow $}{} &= \frac{i}{k^2} (P_V^{-1})^{ab}(P_W^{-1})^{\mu\nu}
\\
\cubicFrule{$a,\mu$}{$b,\nu$}{$c,\rho$}{}{}{} &= i V^{abc}W^{\mu \nu \rho}
\end{split}
\ee
We have introduced $P_V^{ab}$ and $P_W^{\mu\nu}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction, as is the case for $DF^2$ theory \cite{Johansson:2017srf}. Generally, these quadratic dressings are simply delta functions, $P_V^{ab} = \delta^{ab}$, or flat space metrics, $P_W^{\mu\nu}=\eta^{\mu\nu}$. As written, the theory will be color dual if the off-shell cubic vertex is antisymmetric and satisfies the Jacobi identity. Specifically,
\begin{align}
\text{antisymmetry}:& \quad{}_{\mu_1}\langle W^{\mu_2} \rangle_{\mu_3}+\text{cyc}(23)=0
\\
\text{Jacobi identity}:& \quad {}_{\mu_1}\langle W^{\mu_2} W^{\mu_3}\rangle_{\mu_4}+\text{cyc}(234)=0\label{jacID}
\end{align}
where, using the Feynman rules of \cref{CKFrules}, the bracketed expression is the half-ladder numerator of the following diagram,
\be
{}_{\mu_1}\langle W^{\mu_2}W^{\mu_3}\cdots W^{\mu_{n-1}} \rangle_{\mu_n} = N\left[\halfLadderKinetic\right]\,,
\ee
and likewise for the $V^{abc}$ dressing.
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically, when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional operators \cite{Carrasco:2022lbm,Carrasco:2022sck} or propagating fields \cite{Ben-Shahar:2022ixa}.
Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic two-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the loop-level
amplitude integrand must be constructed from an ansatz.  The four
conditions we impose on an integrand ansatz are summarized here and
elaborated on below.\footnote{The four ansatz conditions are similar
  to those of Ref.~\cite{Bern:2015ooa} except that there the authors
  demand that the numerators have the same loop power counting as
  Feynman gauge Feynman rules.  This is an additional assumption that
  is most emphatically not made here.}
After imposing the four constraints below, any remaining coefficients in the ansatz encode ``generalized gauge freedom'', meaning that the
choice of coefficients does not affect the physical integrand or its double copy.
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: A diagram's numerator is only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity}: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

@@

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we argued above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge. We can construct this Lagrangian from \cref{eq:ym-lag} by keeping the right field strength covariant under
$U(N)$, and making the left field strength gauge covariant under
$U(1)^{N^2}$. Note that since $U(1)$ covariance is identical to $U(1)$ invariance, the amplitudes of this semi-abelian theory vanish under $\bar{\varepsilon}(k) \rightarrow k$, where $\bar{\varepsilon}$ is the polarization of an external $\bar{A}_\mu$ vector.

@@

As noted above, we can select out the manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e., SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following sum over building blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog},
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop,
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a\,,
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of
$(\varepsilon k)$ and $(kk)$. As we describe in detail in
\cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to
extract the $D$-dependent part that corresponds to an internal
$\bar{Y}Y$-loop from extra-dimensional scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
$Y$-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

@@

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the relevant BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0,
\end{equation}
that is, the theory is one-loop exact.  In order to produce
non-vanishing interactions, we would need to reintroduce
$D$-dimensional vertices from the full Yang-Mills Lagrangian of
\cref{eq:ym-lag} that we dropped in our construction semi-abelian
YM. In terms of cubic graphs, the additional interaction must
necessarily have the opposite number of $\bar{A}_\mu$ and unbarred
$A_\mu$ fields to that of \cref{semiYMvert}.  Reintroducing these
oppositely oriented vertices allows for new internal contractions of
$\bar{A}_\mu A^\mu$,
\begin{equation}\label{2loopTension}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of
weight\footnote{We choose the convention that the abelianized gauge
  fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian
  vectors are $\mathcal{W}[\bar{A}] = +1$.}
$\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM
vertex which always carries weight $\mathcal{W}[\bar{A}A A] =
+1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. Therefore, by a simple weight counting argument, one can see that
including this wrong sign interactions is unavoidable at two-loop and
higher.

Thus, to construct two-loop numerators prescriptively, as we have done at one-loop, would require knowledge of
the full kinematic algebra for Yang-Mills off-shell.  Since this is presently unavailable, we will tackle the two-loop integrand using an ansatz approach.

>>>>

\abstract{We report on progress in understanding how to construct color-dual multi-loop amplitudes.
First we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators through one-loop directly from Feynman rules. By a simple weight counting argument, this approach does not further generalize to two-loops.  As a first step in
  understanding the two-loop challenge, we use a $D$-dimensional color-dual bootstrap to 
  successfully construct globally color-dual
  \emph{local} two-loop four-point nonlinear sigma model (NLSM) numerators.  The double-copy of
these NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known unitarity constructed integrands 
  of special Galileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively. 
  Applying our bootstrap to two-loop four-point pure Yang-Mills, we exhaustively search the space of local numerators and find that it fails to satisfy global color-kinematics duality, completing a search previously initiated in the literature. 
  We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards non-local
  construction of color-dual integrands at generic loop order.

@@

\section{Introduction}\label{intro}
Since the turn of the century, our understanding of the $S$-matrix and
its concealed structures has expanded dramatically. At the heart of this
progress is the mantra that \emph{physical observables are simpler when
studied on-shell}
\cite{Parke:1986gb,TasiLance,Cheung:2017pzi}. However, while inserting on-shell states certainly offers a path towards taming the
factorial growth of Feynman diagrams, it also obscures the off-shell
simplicity at the heart of many quantum field theories. A shining
example of this hidden structure is the duality between color and
kinematics \cite{BCJ,Bern:2010ue,BCJreview}, which states that the
kinematic numerators of gauge theory amplitudes can be rearranged to
obey the same algebraic identities as the constituent color
factors. When this duality is realized globally off-shell, it
dramatically decreases the combinatorial complexity posed by integrand
construction, and offers a path towards efficient assembly of quantum
gravity integrands directly from simpler gauge theory building blocks
via the double copy \cite{BCJ,Bern:2010ue}.

@@

Many of the one-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}\cdots T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...,\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,..., a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}\cdots f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, ..., a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...,\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed.  Since the sum only contains $(n-2)!$
terms, there must be additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental Bern-Carrasco-Johansson (BCJ) identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+\cdots+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half-ladder because any
diagram can be reduced to a half-ladder through successive
applications of the Jacobi identity.  At one-loop, every diagram can
be related to one topology as well, in this case the $n$-gon, which
can be understood as a forward limit of the half-ladder.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory. We
refer the reader to Ref.~\cite{BCJreview} for more background on the
topic.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to trivialize
the duality by encoding it in an off-shell Lagrangian
of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}\label{cubicCKLag}
\mathcal{L} = P_V^{ab}P_W^{\mu\nu}\mathcal{O}^a_\mu \Box \mathcal{O}^b_\nu +  V^{abc} W^{\mu \nu \rho}\mathcal{O}^a_\mu \mathcal{O}^b_\nu \mathcal{O}^c_\rho
\end{equation}
where $V^{abc}$ and $W^{\mu \nu \rho}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_\mu$ indexed by quantum numbers $a$ and
$\mu$. The Feynman rules for this theory are simply:
\be\label{CKFrules}
\begin{split}
\propFrule{$a,\mu$}{$b,\nu$}{$k \rightarrow $}{} &= \frac{i}{k^2} (P_V^{-1})^{ab}(P_W^{-1})^{\mu\nu}
\\
\cubicFrule{$a,\mu$}{$b,\nu$}{$c,\rho$}{}{}{} &= i V^{abc}W^{\mu \nu \rho}
\end{split}
\ee
We have introduced $P_V^{ab}$ and $P_W^{\mu\nu}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction, as is the case for $DF^2$ theory \cite{Johansson:2017srf}. Generally, these quadratic dressings are simply delta functions, $P_V^{ab} = \delta^{ab}$, or flat space metrics, $P_W^{\mu\nu}=\eta^{\mu\nu}$. As written, the theory will be color dual if the off-shell cubic vertex is antisymmetric and satisfies the Jacobi identity. Specifically,
\begin{align}
\text{antisymmetry}:& \quad{}_{\mu_1}\langle W^{\mu_2} \rangle_{\mu_3}+\text{cyc}(23)=0
\\
\text{Jacobi identity}:& \quad {}_{\mu_1}\langle W^{\mu_2} W^{\mu_3}\rangle_{\mu_4}+\text{cyc}(234)=0\label{jacID}
\end{align}
where, using the Feynman rules of \cref{CKFrules}, the bracketed expression is the half-ladder numerator of the following diagram,
\be
{}_{\mu_1}\langle W^{\mu_2}W^{\mu_3}\cdots W^{\mu_{n-1}} \rangle_{\mu_n} = N\left[\halfLadderKinetic\right]\,,
\ee
and likewise for the $V^{abc}$ dressing.
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically, when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional operators \cite{Carrasco:2022lbm,Carrasco:2022sck,Carrasco:2023wib} or propagating fields \cite{Ben-Shahar:2022ixa}.
Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic two-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the loop-level
amplitude integrand must be constructed from an ansatz.  The four
conditions we impose on an integrand ansatz are summarized here and
elaborated on below.\footnote{The four ansatz conditions are similar
  to those of Ref.~\cite{Bern:2015ooa}. However, in that work the authors
  demand that the numerators have the same loop power counting as Feynman rules for vector fields in Feynman gauge.  This additional assumption of theirs, which most emphatically is not made here, places more restrictive constraints on the ansatz than our conditions.}
After imposing the four constraints below, any remaining coefficients in the ansatz encode ``generalized gauge freedom'', meaning that the
choice of coefficients does not affect the physical integrand or its double copy.
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: A diagram's numerator is only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity}: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

@@

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we argued above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge, so that $\partial_\mu A^\mu =0$. We can construct this Lagrangian from \cref{eq:ym-lag} by keeping the right field strength covariant under
$U(N)$, and making the left field strength gauge covariant under
$U(1)^{N^2}$. Note that since $U(1)$ covariance is identical to $U(1)$ invariance, the amplitudes of this semi-abelian theory vanish under $\bar{\varepsilon}(k) \rightarrow k$, where $\bar{\varepsilon}$ is the polarization of an external $\bar{A}_\mu$ vector.

@@

As noted above, we can select out the manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e., SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following sum over building blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog},
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop,
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a\,,
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of
$(\varepsilon k)$ and $(kk)$. As we describe in detail in
\cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to
extract the $D$-dependent part that corresponds to an internal
$\bar{Y}Y$-loop from extra-dimensional scalars,
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
$Y$-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

@@

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the relevant BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0,
\end{equation}
that is, the theory is one-loop exact.  In order to produce
non-vanishing interactions, we would need to reintroduce
$D$-dimensional vertices from the full Yang-Mills Lagrangian of
\cref{eq:ym-lag} that we dropped in our construction semi-abelian
YM. In terms of cubic graphs, the additional interaction must
necessarily have the opposite number of $\bar{A}_\mu$ and unbarred
$A_\mu$ fields to that of \cref{semiYMvert}.  Reintroducing these
oppositely oriented vertices allows for new internal contractions of
$\bar{A}_\mu A^\mu$,
\begin{equation}\label{2loopTension}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of
weight\footnote{We choose the convention that the abelianized gauge
  fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian
  vectors are $\mathcal{W}[\bar{A}] = +1$.}
$\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM
vertex which always carries weight $\mathcal{W}[\bar{A}A A] =
+1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. Thus, by a simple weight counting argument, one can see that
including this wrong sign interactions is unavoidable at two-loop and
higher.

Therefore, to construct two-loop numerators prescriptively, as we have done at one-loop, would require knowledge of
the full kinematic algebra for Yang-Mills off-shell.  Since this is presently unavailable, we will tackle the two-loop integrand using an ansatz approach.
